# Updating CRAN Mirror

Before we install the packages, let's make sure we are pulling from a recent mirror of CRAN. By default MRS pulls from an old version of CRAN, usually around the time that version of MRS was released.

```{r update_repos}

r <- getOption('repos')
# set mirror to something a bit more recent
r["CRAN"] <- "https://mran.revolutionanalytics.com/snapshot/2016-07-05/"
options(repos = r)


```


# Install magrittr and ggplot2 

We'll install some familiar packages

```{r gg_pipe}

install.packages('magrittr')
install.packages('ggplot2')


```



# Install curl and gnutls for devtools

We'll need a couple of packages that are not yet avaialable on CRAN, but are available as a development release on github.

```{r devtools_install}

system("sudo apt-get -y build-dep libcurl4-gnutls-dev")
system("sudo apt-get -y install libcurl4-gnutls-dev")
install.packages('devtools')


```



# Install sparkapi and sparklyr packages 

```{r git_install}

devtools::install_github("rstudio/sparkapi")
devtools::install_github("rstudio/sparklyr")
devtools::install_github('ropensci/plotly')

```

