# Updating CRAN Mirror

Before we install the packages, let's make sure we are pulling from a recent mirror of CRAN. By default MRS pulls from an old version of CRAN, usually around the time that version of MRS was released.

```{r update_repos}

r <- getOption('repos')
# set mirror to something a bit more recent
r["CRAN"] <- "https://mran.revolutionanalytics.com/snapshot/2016-12-19/"
options(repos = r)


```


# Install magrittr and ggplot2 

We'll install some familiar packages

```{r gg_pipe}

install.packages('magrittr')
install.packages('ggplot2')


```



# Install curl and gnutls for devtools

We'll need a couple of packages that are not yet avaialable on CRAN, but are available as a development release on github..

```{r devtools_install}
# 
# system("sudo apt-get -y build-dep libcurl4-gnutls-dev")
# system("sudo apt-get -y install libcurl4-gnutls-dev")
# system("sudo ln -s /bin/tar /bin/gtar")
# install.packages('devtools')

```



# Install sparkapi and sparklyr packages 

```{r git_install}


install.packages('sparklyr')
install.packages("plotly")

# devtools::isntall_github("rstudio/htmltools")
# devtools::install_github("rstudio/rmarkdown")
# devtools::install_github("rstudio/sparklyr")
# devtools::install_github("hadley/purrr")
# devtools::install_github('ropensci/plotly')
# devtools::install_github('yihui/knitr')

```

