# Modeling with RxSpark

This section shows how to use the `RxSpark` compute context for modeling.

# Locate RevoShare dir

Every MRS installation on a HDFS environment creates a share directory on HDFS. By default, each user will have her own shared directory under the `/user/RevoShare/` file path.

```{r revoshare}

rxHadoopListFiles("/user/RevoShare/")
data_path <- "/user/RevoShare/alizaidi"


```

# Saving the Spark DataFrame to CSV

The `RxSpark` and the Spark Compute contexts are completely distinct compute environments. In order to use the `rx` functions, we need to move the Spark DataFrame into a format that MRS can understand.

## Write Sample Taxi to RevoShare 

```{r save_csv}

library(sparklyr)
spark_write_csv(sample_taxi, 
                path = file.path(data_path, 'sampleTaxi'))


```


## Remove SUCCESS dir

```{r remove_success}

rxHadoopListFiles(file.path(data_path, "sampleTaxi"))
file_to_delete <- file.path(data_path, 
                            "sampleTaxi", "_SUCCESS")
delete_command <- paste("fs -rm", file_to_delete)
rxHadoopCommand(delete_command)


```


# Create HDFS and Spark Contexts for Revo

Let's create the pointers to the file paths and HDFS to use the `RxSpark` compute context.

```{r hdfs_pointers}

myNameNode <- "default"
myPort <- 0
hdfsFS <- RxHdfsFileSystem(hostName = myNameNode, 
                           port = myPort)

taxi_text <- RxTextData(file.path(data_path,
                                  "sampleTaxi"),
                        fileSystem = hdfsFS)

taxi_xdf <- RxXdfData(file.path(data_path, "taxiXdf"),
                      fileSystem = hdfsFS)


```




## Import to XDF

Now we use our `rxImport` function to import the csv into an xdf.

```{r csv_import_xdf}

rxImport(inData = taxi_text, taxi_xdf, overwrite = TRUE)
rxGetInfo(taxi_xdf)


```



# create RxSpark compute context

```{r rx_lin_mod}


computeContext <- RxSpark(consoleOutput=TRUE,
                          nameNode=myNameNode,
                          port=myPort,
                          executorCores=6, 
                          executorMem = "3g", 
                          executorOverheadMem = "3g", 
                          persistentRun = TRUE, 
                          extraSparkConfig = "--conf spark.speculation=true")

rxSetComputeContext(computeContext)

taxi_Fxdf <- RxXdfData(file.path(data_path, "taxiXdfFactors"),
                       fileSystem = hdfsFS)


rxFactors(inData = taxi_xdf, outFile = taxi_Fxdf, 
          factorInfo = c("pickup_hour", "pickup_nhood")
)

system.time(linmod <- rxLinMod(tip_pct ~ trip_distance, 
                               data = taxi_xdf, blocksPerRead = 2))


```

